---
title: "Homework 03"
---


## TidyTuesday Section (optional)

::: {.callout-important title="Instructions"}

You can count work on this week's TidyTuesday toward the exceptional work required for an A in the Homework component.

:::


Explore the week's [TidyTuesday](https://github.com/rfordatascience/tidytuesday) challenge. Develop a research question, then answer it through a short data story with [effective visualization(s)]({{< var effective-viz-url >}}).  Provide sufficient background for readers to grasp your narrative.

```{r}
#Necessary Libraries
#| include: false
#| echo: false

library(readxl)
library(dplyr)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(tidytuesdayR)
```

##Importing the Data 

Reading in the Tidytuesday data for last week, which takes a look at recipes gathered Allrecipes.com. The first data set -- all_recipes -- has a collection of regular recipes and the second data set -- cuisines -- contains recipes organized by their country of origin, and has fewer data points than the all_recipies data set. 

```{r}
#| include: true
#| echo: true

tt_data <- tt_load("2025-09-16")
all_recipes <- tt_data$all_recipes
cuisines <- tt_data$cuisines
```

Research Questions: What recipe takes the longest and what is the fastest? Does a certain cuisine take longer than others, how much of it is prep vs cook time? 

## Exploring the Data

```{r}
#Joining the Data 
recipes_expanded <- full_join(all_recipes, cuisines)
```

#Finding recipe with longest total time
```{r}
recipes_expanded |> 
  slice_max(total_time, with_ties = TRUE)
```
The recipe that takes the longest is homemade wine, which funnily enough has only 5 min prep time and 0 cook time but needs to ferment for about 42 days. It does make 32 servings though, probably for a wine tasting party. 

#Finding recipe with shortest total time
```{r}
recipes_expanded |> 
  filter(total_time == 1) #over 300 had cooking times of zero, which was most likely put in place because no time was given, which is why I started with == to 1.
```
The 4 recipies that take the "shortest" amount of time are all sauces/spice mixes. Pumpkin pie spice and 3 ingredient Potsticker dip, do sound like they would take minimal prep time which is why their authors have most likely just marked the time as 1 minute. However teriyaki sauce and sauteed oyster mushrooms in garlic butter sound like they might take more time in reality. Since these are user entered -- volunteer data -- there doens't have to be accuracy for reporting, which is most likely why these recipes show up here. It is important to then understand that user error might be a serious bias skewing or limited analysis. 

#Cusine with longest average time
```{r}
cuisines |> 
  group_by(country) |> 
  summarise(Mean_time = mean(total_time)) |> 
  arrange(desc(Mean_time))
```
It appears Amish and Mennonite food takes the longest to make, most likely because they do things more traditionally and preserve a lot of food which takes time. In comparison on average Colombian food takes the least average amount of time. A lot of the mean times however are quite long because they are most likely being skewed by a few large data points or recipes that take a long time. 

## Visualizations 

#Prepping the Data for the Visualization
```{r}
Mean_times_cuisines <-  cuisines |> 
  mutate(other_time = (total_time - (prep_time + cook_time))) |> 
  group_by(country) |> 
  summarise(mean_prep_time = mean(prep_time), mean_cook_time = mean(cook_time), mean_other_time = mean(other_time)) |> 
  pivot_longer(
    cols = c(mean_prep_time, mean_cook_time, mean_other_time),
    names_to = "work_type", 
    values_to = "time_dedicated"
  )
```

#Vizualization 
```{r}
Mean_times_cuisines |> 
  ggplot(aes(x = "", y = time_dedicated, group = work_type, colour = work_type, fill = work_type)) + 
  geom_bar(width = 1, stat = "identity") +
  coord_polar("y", start=0) + 
  facet_grid(.~country) +
  theme_void()
```


